---
title: "Data Analysis Project"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE,
                      comment = "")
```



# Student Information

**Name**: Sahithi Oddiraju
**G Number**: G01160714




# Introduction

This semester we will be working with a dataset of auto claims filed by customers of an automobile insurance company located in the southwest and western regions of the United States.

Insurance companies depend on accurate pricing estimates to maintain profitability. Auto policies must be priced so that the insurance company makes a profit in the long run, given the costs of their customers' payouts for accident repairs, total loss car replacements, medical fees, and legal fees.

The executives at this insurance company have noticed declining profitability over the last several years and have hired you as a data science consultant to evaluate their claims data and make recommendations on pricing, customer behavior, and car insurance policy adjustments.

The objective of this project is to perform an exploratory data analysis on the claims_df dataset and produce an executive summary of your key insights and recommendations to the executive team at the insurance company.

Before you begin, take a moment to read through the following insurance company terms to familiarize yourself with the industry:  [Auto Insurance
Terms](https://www.iii.org/article/auto-insurance-jargon-buster "Auto Insurance Terms"){target="_blank"} 



## Auto Claims Data

The `claims_df` data frame is loaded below and consists of 6,249 auto claims submitted by customers of the insurance company. The rows in this data frame represent a single claim with all of the associated features that are displayed in the table below.

**Note**: If you have not installed the `tidyverse` package, please do so by going to the `Packages` tab in the lower right section of RStudio, select the `Install` button and type `tidyverse` into the prompt. If you cannot load the data, then try downloading the latest version of R (at least 4.0). The `readRDS()` function has different behavior in older versions of `R` and may cause loading issues.


```{r}
library(tidyverse)

claims_df <- readRDS(url('https://gmubusinessanalytics.netlify.app/data/claims_df.rds'))

```


### Raw Data

```{r}

claims_df

```
<br>
<br>


# Exploratory Data Analysis


Executives at this company have hired you as a data science consultant to evaluate their claims data and make recommendations on pricing, customer behavior, and car insurance policy adjustments.

You must think of **at least 8 relevant questions** that will provide evidence for your recommendations.

The goal of your analysis should be discovering which variables drive the differences between customers with large lifetime values and
customers who cost the company more than they provide in revenue through monthly premiums.

Some of the many questions you can explore include:

-   Are there types of customers, based on their policy or demographics,
    that are highly profitable?

-   Do certain policies have a lower number of claims, leading to large
    profits?

-   Are there "problem customers" which have a large number of claims?

You must answer each question and provide supporting data summaries with either a summary data frame (using `dplyr`/`tidyr`) or a plot (using `ggplot`) or both.

In total, you must have a minimum of 5 plots and 4 summary data frames for the exploratory data analysis section. Among the plots you produce, you must have at least 4 different types (ex. box plot, bar chart, histogram, heat map, etc...)

Each question must be answered with supporting evidence from your tables and plots.





## Question 1



**Question**: What is the vehicle type which pays the most premium



**Answer**: The vehicle that pays the highrsta mount of premium is mid size type.



To add additional R code chunks for your work, select `Insert` then `R` from the top of this notebook file.

```{r}

df_1 <- claims_df %>% group_by(vehicle_size) %>%
        summarize(avg_monthly_premium_vehicle=sum(monthly_premium))
df_1<-arrange(df_1, desc(avg_monthly_premium_vehicle))
df_1
ggplot(data = df_1, aes(x=vehicle_size, y= avg_monthly_premium_vehicle)) +
  geom_bar(stat = "identity", fill="Red", alpha=0.6, width=0.4) +
  ggtitle("Size of Vehicle vs Premium Amount Paid") + 
  theme(plot.title = element_text(hjust = 0.5))+ 
  labs(x="Size Of Vehicle",y = "Premium amount")+
  theme(axis.text.x = element_text(angle = 90))


```



## Question 2



**Question**:What is the state with highest claims



**Answer**: California is the state with highest claims



```{r}
claims_df
ggplot(claims_df, 
       aes(x = customer_state, 
           fill = total_claims)) + 
  geom_bar(position = "dodge",fill="orange")+
  ggtitle(" customer_state vs total_claims ")
  
```



## Question 3



**Question**: In which state are poeple having more claims greater than 3 and 4?



**Answer**: California is the state which has highest number of claims in the category where number of claims is 3 and no of claims is 4




```{r}

df_3 <- claims_df%>% filter(total_claims>=3)
df_3$total_claims<- as.factor(df_3$total_claims)
typeof(df_3$total_claims)
plot3<-ggplot(data = df_3, aes(x=customer_state,fill=total_claims))+
  geom_bar(stat = "count", position="dodge", alpha=0.6, width=0.4) +
  ggtitle(" customer_state vs total_claims ") + 
  theme(plot.title = element_text(hjust = 0.5))+ 
  labs(x="customer_state",y = "total_claims")+
  theme(axis.text.x = element_text(angle = 20))

plot3





```




## Question 4



**Question**: Distribution of residences among claims



**Answer**:people living in residence  type sub urban are the highest number of people opting for policies



```{r}
df_3
df_3$total_claims <- as.factor(df_3$total_claims)

plot4<-ggplot(data = df_3, aes(x=residence_type,fill=total_claims))+
  geom_bar(stat = "count", position="dodge", alpha=0.6, width=0.4) +
  ggtitle("Relationship between Residence and Policies ") + 
  theme(plot.title = element_text(hjust = 0.5))+ 
  labs(x="Residence",y = "Policy count")
  

plot4
```




## Question 5



**Question**: What is the highest type of policy type opted 



**Answer**: Personal is the highest opted policy type




```{r}

claims_df$total_claims<- as.factor(claims_df$total_claims)
ggplot(df_3, aes(x=current_claim_amount, y=months_since_last_claim,color=total_claims)) + 
  geom_point(aes(size=total_claims))


```


## Question 6



**Question**:Relationship between policy type and premium amount



**Answer**: From the previous question we found out that, the most people are opting 
for personal policy type. here the premium amount is also the highest for the personal policy type.
Hence the policy type and premium amount are positively correlated.




```{r}
df_6 <- df_3 %>% filter(total_claims==4) %>% filter(current_claim_amount> 2000)
df_6
p <- ggplot(df_6, aes(x=vehicle_size, y=current_claim_amount)) + 
  geom_boxplot()
p


```


## Question 7



**Question**: What is the highest state opting for personal policy types



**Answer**: California is state with 1576 people opting for personal policy type




```{r}

df_6 <- df_6 %>% filter(vehicle_size=="Midsize")
df_6
plot6<-ggplot(data = df_6, aes(x=customer_state,y=current_claim_amount,fill=customer_state))+
  geom_bar(stat = "identity", position="dodge", alpha=0.6, width=0.4) +
  ggtitle("Relationship between customer_state and current_claim_amount ") + 
  theme(plot.title = element_text(hjust = 0.5))+ 
  labs(x="customer_state",y = "current_claim_amount")+
  theme(axis.text.x = element_text(angle = 20))

plot6

```



## Question 8



**Question**:Most Policy type opted by Suburban people



**Answer**: As per statistics, people in sub urban areas are opting for Personal policy type.




```{r}
df_7 <- df_6 %>% filter(customer_state=="Oregon")
df_7
df_7<-df_7 %>% select(residence_type,marital_status,coverage,policy,vehicle_class)
summary(df_7)


```


# Summary of Results

Write an executive summary of your overall findings and recommendations to the executives at this company. Think of this section as your closing remarks of a presentation, where you summarize your key findings and make recommendations to improve pricing, company operations, and car insurance policy adjustments.

Your executive summary must be written in a [professional tone](https://www.universalclass.com/articles/writing/business-writing/appropriate-tone-in-business-communications.htm){target="_blank"}, with minimal grammatical errors, and should include the following sections:

1.  An introduction where you explain the business problem and goals of your data analysis

    -   What problem(s) is this company trying to solve? Why are they
        important to their future success?
        

    -   What was the goal of your analysis? What questions were you
        trying to answer and why do they matter?

<br>
Since this insurance company's executives have observed declining profitability over the past several years, they want to review their claims data and give suggestions regarding pricing, consumer behavior, and modifications to vehicle insurance policies.
This project's goal is to conduct an exploratory data analysis on the claims df data set and then prepare an executive summary of  key findings and suggestions for the insurance company's executive management.

2.  Highlights and key findings from your Exploratory Data Analysis section

    -   What were the interesting findings from your analysis and **why
        are they important for the business**?

    -   This section is meant to **establish the need for your
        recommendations** in the following section
        

<br>
• Mid size Vehicles are the vehicles which pay the highest amount of premium
• California is the state with highest number of claims
• Since the company is facing issues with the claims the interest here would be find out which state has the count with highest number of claims is greater than 3,4. Again California is the state which has highest number of claims in number of claims =3, and number of claims=4
• The customers living in suburban areas have made highest amount of claims
• There are lot of customers who have claimed more than 4 times and the amount is also greater than 2000 dollars.
• Midsize vehicles are the vehicles with highest number of claims with number of vehicles greater than 4 and current claim greater than 2000
• Oregon is the state with highest number of claims
• In Oregon state the customers with highest claims are opting for a personal policy

3.  Your recommendations to the company

    -   Each recommendation must be supported by your data analysis
        results

    -   You must clearly explain **why** you are making each
        recommendation and which results from your data analysis support
        this recommendation

    -   You must also describe the potential business impact of your
        recommendation:

        -   Why is this a good recommendation?

        -   What benefits will the business achieve?
Based on the analysis, customers from California have the highest number of claims and the 
mid size vehicles are the source of revenue as well as the source of expenditure for the company.
Most of the midsize vehicles are claiming amount greater than 2000 in less than 10 months in the state of oregon. And the policy type opted is personal with basic coverage.
From the data, it is suggested to make changes to the insurance plan of midsize vehicles personal plan.


## Executive Summary

Please write your executive summary below. If you prefer, you can type your summary in a text editor, such as Microsoft Word, and paste your final text here.

When summarized , the most number of claims are made by mid size vehicles 
from Oregon and California.
The policy type of such claims chosen is the Personal policy type with basic
coverage.
Highest amount of claims are made by customers from sub urban area.













